### 2018-05-11 slides/pnwplse.tex
more work on talk

 \documentclass{beamer}
\usefonttheme{professionalfonts}
\usepackage{agda}
\usepackage{catchfilebetweentags}
\usetheme{CambridgeUS}
\usecolortheme{seahorse}

\usepackage{amssymb}
\usepackage{bbm}
\usepackage[greek,english]{babel}

\beamertemplatenavigationsymbolsempty

\title{Musical Ornaments}
\author{John Leo}
\institute{Halfaya Research}
\date{May 14, 2018}

\begin{document}

\frame{\titlepage}

\begin{frame}\frametitle{Outline}
\begin{itemize}
\item A look toward the future
\item Music Tools
\item Equivalences
\item Ornaments
\end{itemize}

\bigskip 
Sources:
\begin{itemize}
\item {\tt https://github.com/halfaya/MusicTools}
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Robert Harper}
\begin{quote}
Eventually all the arbitrary programming languages are going to be just swept away with the oceans,
and we will have the permanence of constructive, intuistionistic type theory as the master theory
of computation---without doubt, in my mind, no question.  So, from my point of view---this is a personal
statement---working in anything else is a waste of time.
\end{quote}

CMU Homotopy Type Theory lecture 1, 52:56--53:20.
\end{frame}

\begin{frame}\frametitle{What will programming look like in 50 years?}
\begin{itemize}
\item Convergence of math and computer science
\item Functional Programming, Algebra of Programming
\item Dependent Types or a successor (Cubical?)
\item Who does the programming?
\end{itemize}
\end{frame}

\begin{frame}\frametitle{How do we get there from here?}
\begin{itemize}
\item Add dependent types to an industrial-strength language (Haskell)
\item Make a dependently typed language (Agda, Idris) practical to use
\item Learn how to program using dependent types
\item Many theoretical and practical advances are still needed
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Euterpea}
\begin{figure}
\includegraphics[height=0.8\textheight]{hsom}
\end{figure}
\end{frame}

\begin{frame}\frametitle{Look vs Time (1997)}
\begin{figure}
\includegraphics[width=\textwidth]{lookVsTime}
\end{figure}
\end{frame}

\begin{frame}\frametitle{Music Representation \`a la Euterpea}
\ExecuteMetaData[../agda/latex/Pnwplse.tex]{pitch}
\ExecuteMetaData[../agda/latex/Pnwplse.tex]{duration}
\ExecuteMetaData[../agda/latex/Pnwplse.tex]{note}
\ExecuteMetaData[../agda/latex/Pnwplse.tex]{music}
\end{frame}

\begin{frame}\frametitle{Conclusion}
Goes here.
\end{frame}

\end{document}