### 2019-06-24 doc/farm/farm-abstract.tex
minor fixes

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[sigplan,10pt,screen]{acmart}
\settopmatter{printfolios=true, printccs=true, printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

\acmConference[FARM 2019]{The ACM SIGPLAN International Workshop on Functional Art, Music, Modelling and Design}{August, 2019}{Berlin, Germany}
\acmYear{2019}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{alltt}

\begin{document}

%% Title information
\title{Demo: Counterpoint by Construction}
%\subtitle{Functional Pearl}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.

%% Author with single affiliation.
\author{Youyou Cong}
\affiliation{
  \institution{Tokyo Institute of Technology}            %% \institution is required
  \city{Tokyo}
  \country{Japan}
}
\email{cong@c.titech.ac.jp}          %% \email is recommended

\author{John Leo}
\affiliation{
  \institution{Halfaya Research}            %% \institution is required
  \city{Bellevue}
  \state{WA}
  \country{USA}
}
\email{leo@halfaya.org}          %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
We present \emph{Music Tools}, an Agda library for analyzing 
and synthesizing music.  
The library uses dependent types to simplify encoding of 
music rules, thus improving existing approaches based on 
simply typed languages.  
As an application of the library, we demonstrate an 
implementation of first-species counterpoint, where we use
dependent types to constrain the motion of two parallely
sounding voices.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010405.10010469.10010475</concept_id>
<concept_desc>Applied computing~Sound and music computing</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
<concept_desc>Software and its engineering~Functional languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Sound and music computing}
\ccsdesc[500]{Software and its engineering~Functional languages}
%% End of generated code

%% Keywords
%% comma separated list
% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

Western music of the common practice period tends to loosely
follow sets of rules, which were developed over time to ensure
the aesthetic quality of the composition. 
Among such rules, those for harmony \citep{piston1987harmony}
and counterpoint (harmonically interdependent melodies) 
\citep{fux1965study} are particularly fundamental and continue 
to be taught to music students, not only as a means to understand 
the music of that period, but also as a foundation for modern art 
and popular music.

To analyze and synthesize tonal music, researchers have
attempted to encode these rules into programming languages.
Functional programming languages seem
ideally suited for this task, and in particular, those with a static
type system can further guarantee that \emph{well-typed music 
does not sound wrong}.
In the past decade, Haskell has been extensively used to encode
the rules of harmony \citep{fmmh, hihseufha, faamh, hafha, fghm}
as well as counterpoint \citep{szamozvancev2017well}.
An interesting observation is that, many of the existing encodings
rely on some form of \emph{dependent types}, i.e.,  
types that depend on terms. 
Since Haskell is not a dependently typed language, one has to
use language extensions, such as GADTs \citep{cheney2002lightweight} 
and singleton types \citep{eisenberg2013dependently},
to simulate dependencies.
While this allows encoding a wide class of music rules, it can
require duplicating code to reflect terms into 
types, making the implementation less elegant 
\citep{monnier2010singleton}.
This motivates us to explore music programming in a language
with intrinsic support for dependent types.

We present Music Tools\footnote{\url{https://github.com/halfaya/MusicTools}},
a library of small tools that can be combined functionally to 
help analyze and synthesize music.
To allow simple and natural encoding of rules, we build our 
library in Agda \citep{norellphd}, a functional language with full dependent types.
As an application of the library, we demonstrate an
implementation  of species counterpoint, based on the rules 
given by \citet{fux1965study}.
Thanks to Agda's rich type system, we can express 
the rules in a straightforward manner, and thus ensure by 
construction that well-typed counterpoint satisfies all the 
required rules.

\section{The Music Tools Library}

TODO: describe the key ingredients of the library.

\section{Application: First-Species Counterpoint}

We now explain how to implement the rule system of 
first-species counterpoint\footnote{The code is available at \\
\url{https://github.com/halfaya/MusicTools/blob/master/agda/Counterpoint.agda}.}.
In first-species counterpoint, one starts with a base melody 
(the \emph{cantus firmus}), and constructs a counterpoint 
melody note-by-note in the same rhythm.
The two voices are represented as a list of pitch-interval pairs,
where intervals must not be dissonant (2nds, 7ths, or 4ths).

\begin{alltt}
data IntervalQuality : Set where
  min3  : IntervalQuality
  maj3  : IntervalQuality
  per5  : IntervalQuality
  min6  : IntervalQuality
  maj6  : IntervalQuality
  per8  : IntervalQuality
  min10 : IntervalQuality
  maj10 : IntervalQuality

PitchInterval : Set
PitchInterval = Pitch \(\times\) IntervalQuality
\end{alltt}

In addition, it is prohibited to move from any interval to
a perfect interval (5th or octave) via parallel or similar motion.
Therefore, we define a predicate that checks whether a motion 
is allowed or not.

\begin{alltt}
motionOk : (i1 : Interval)
           (i2 : Interval) \(\rightarrow\) Set
motionOk i1 i2 with motion i1 i2
         | isPerfectInterval i2
motionOk i1 i2 | contrary | \_     = \(\top\)
motionOk i1 i2 | oblique  | \_     = \(\top\)
motionOk i1 i2 | parallel | false = \(\top\)
motionOk i1 i2 | parallel | true  = \(\bot\)
motionOk i1 i2 | similar  | false = \(\top\)
motionOk i1 i2 | similar  | true  = \(\bot\)
\end{alltt}

The last requirement is that the music must end with a cadence,
which is a final motion from the 2nd or 7th degree to the tonic 
(1st degree). 
We impose this requirement by declaring two cadence constructors 
as the base cases of counterpoint.
Thus, we arrive at the following datatype for well-typed counterpoint\footnote{
For readability, we have omitted explicit conversions 
from \texttt{PitchInterval} (which ensures the interval is not dissonant) 
to the general \texttt{Interval}.}.

\begin{alltt}
data FirstSpecies : PitchInterval \(\rightarrow\)
                    Set where
  cadence2 : (p : Pitch) \(\rightarrow\)
    FirstSpecies (transpose (+ 2) p , maj6)
  cadence7 : (p : Pitch) \(\rightarrow\)
    FirstSpecies (transpose -[1+ 0 ] p , min10)
  \_::\_ : (pi : PitchInterval) \(\rightarrow\)
         \{pj : PitchInterval\} \(\rightarrow\)
         \{\_ : motionOk pi pj\} \(\rightarrow\)
         FirstSpecies pj \(\rightarrow\)
         FirstSpecies pi
\end{alltt}

\noindent Observe that \texttt{motionOk} is an implicit argument of 
the \texttt{\_::\_} constructor. 
The argument can be resolved automatically by the type checker,
hence there is no need to manually supply this proof.

Now we can write valid first-species counterpoint as in the 
example below.

\begin{alltt}
example : FirstSpecies (g 4 , per8)
example = 
  (g 4 , per8) :: (c 5 , maj10) ::
  (c 5 , per8) :: (c 5 , maj10) ::
  (e 5 , min10) :: (g 5 , per8) ::
  (cadence2 (c 6))
\end{alltt}

\section{Future Work}

TODO

%At the FARM workshop, we intend to give a gentle introduction to
%counterpoint, and describe our Agda implementation, showing how the
%type-based approach both aids human composition and allows for 
%computer-generated creation of correct counterpoint.
%We then contrast our work to a recent study on generating
%natural-sounding counterpoint by machine learning
%\citep{CounterpointByConvolution}, which does not provide correctness
%guarantees.
%Finally, we discuss further applications of our library, including
%representation of functional harmony.

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
 The authors would like to thank the participants of the Tokyo Agda 
 Implementors' Meeting, especially Ulf Norell and Jesper Cockx,
 for many helpful suggestions that improved our Agda code.
\end{acks}

%% Bibliography
\bibliography{farm-abstract.bib}

\end{document}