### 2020-02-20 doc/icfp20/conclusion.tex
start introduction

\section{Conclusion}
\label{sec:conclusion}

Text here.