### 2020-02-29 doc/icfp20/conclusion.tex
wip

\section{Conclusion}
\label{sec:conclusion}

TODO: easy to express rules, break into parts, reuse in other contexts, extend.
Downsides: hard to do some things (equality, sets, etc), Proof emphasis can get in the way, but
may want to use in the future. Future work: equivalence, ornaments, ML.