### 2020-03-04 doc/icfp20/harmony.tex
major work on harmony section

\newcommand{\Harmonization}{
  \begin{figure}
    \includegraphics[width=5.5cm]{fig/piston.png}
    \hspace{1cm}
    \includegraphics[width=5cm]{fig/fharm.png} \\
    \begin{flushleft}
    \begin{small}
     \hspace{2.35cm} V \ \ \ \ \ \ \ \,I \ \ \ \ \ \ \ \,IV \,VI \ \,III \ IV \ \ \,V
     \hspace{2.6cm} V \ \ \ \ \ \ \ \,I \ \ \ \ \ \,IV VI \,III\,IV \ \ V
    \end{small}
    \end{flushleft}
    \caption{Harmonization by Piston (left) and \fharm (right)}
    \Description{Harmonization by Piston (left) and \fharm (right)}
    \label{fig:harmonization}
  \end{figure}
}

\newcommand{\HarmonizationMT}{
  \begin{figure}
    \includegraphics[width=5.5cm]{fig/piston.png}
    \hspace{1cm}
    \includegraphics[width=5cm]{fig/fharm.png} \\
    \begin{flushleft}
    \begin{small}
     \hspace{2.35cm} V \ \ \ \ \ \ \ \,I \ \ \ \ \ \ \ \,IV \,VI \ \,III \ IV \ \ \,V
     \hspace{2.6cm} V \ \ \ \ \ \ \ \,I \ \ \ \ \ \,IV VI \,III\,IV \ \ V
    \end{small}
    \end{flushleft}
    \caption{Harmonization by Music Tools (to be replaced by real figure)}
    \Description{Harmonization by Music Tools}
    \label{fig:harmonizationMT}
  \end{figure}
}

\section{Harmony}
\label{sec:harmony}

Harmony refers to simultaneously sounding tones (chords) and its study
is primarily concerned with the progression of chords throughout a
piece of music. It is in essence the of dual of counterpoint. In this
section we formalize a small part of the classic text
\citet{piston-harmony}, in particular part of Chapter~9 which is
concerned with techniques for harmonizing a given melody.

\subsection{Harmonic Progression}
\label{sec:harmony:prog}

The primary Example~9.1 in Piston's text is shown on the left
side of Figure~\ref{fig:harmonization} which shows the melody with a
sample four part SATB harmonization in C~major. The melody is the
series of highest notes, the soprano (S) line, while the other four
parts alto (A), tenor (T) and bass (B) are the melodies below
it. Roman numerals denote the names of the triads at each vertical
slice. Note that even though there are four voices there are only
three distinct pitches sounding at each vertical slice (where tones an
octave apart are considered to have the same relative pitch). As is
typical the root of the triad is always doubled. Also note that all
triads are in root position (meaning the root is in the bass voice),
as Piston restricts to this case initially in the chapter (he does
relax it later). The goal of a harmonization is both to create an
aesthetically pleasing chord progression as well as fluid melodies in
each voice (called voice leading) and independence of each music
line (in other words good counterpoint). In this chapter as well as
throughout the text Piston gives advice on how to achieve these goals.

\Harmonization

We would like to formalize some of this advice and apply it to
automate harmonization of melodies. We start with selection of the
sequence of triads, in other words the harmonic progression. Earlier
in the text on page~23 Piston presents a Table of Usual Root
Prrogressions, noting which triads usually follow a given one and
which are sometimes used or rare. It is easy to encode this in Agda.

\begin{alltt}
data Triad : Set where
  I : Triad; II : Triad; III : Triad; IV : Triad; V : Triad; VI : Triad; VII : Triad

record NextTriad : Set where
  constructor nextTriadb
  field
    usual : List Triad; sometimes : List Triad; rare : List Triad

rootProgression : Triad \(\rightarrow\) NextTriad
rootProgression I   = nextTriad (IV :: V :: []) (VI :: []) (II :: III :: [])
rootProgression II  = nextTriad (V :: []) (IV  :: VI :: []) (I :: III :: [])
-- other cases omitted for space
\end{alltt}

It is easiest to create a harmonic progression in reverse, to
ensure one ends a musical phrase with either a full (V--I) or half (V)
cadence, and work backward from there. So we would like to invert this
table to give a list of likely triads to precede a given triad. This
would normally be straightfoward but here we see a disadvantage of
using Agda. As is typical with languages based on dependent type
theories, equality is a very subtle concept and even determing if a
term is a member of a list is nontrivial.

In Haskell this would be easy--there is an \texttt{Eq} typeclass and a
method for automatically generating equalities. In the Agda Standard
Library~\citep{agda-stdlib} is a relation with constructive proof of
membership, and requires far more effort than we need. Alternatives
include defining decidable equality for triads (which requires writing
out 49 cases) or mapping to natural numbers and using already
implemented decidable or boolean equality for those. For now we simply
write the inverse function manually, excluding the rare cases.

\begin{alltt}
previousTriads : Triad → List Triad
previousTriads I   = V :: IV :: VII :: []
previousTriads II  = VI :: IV :: []
-- other cases omitted for space
\end{alltt}

\subsection{Completing the Harmonization}
\label{sec:harmony:complete}

TODO: Finish this.

See Figure~\ref{fig:harmonizationMT}, which is currently just a copy of
the other figure, but included here now for space.

\HarmonizationMT

\subsection{Comparison with Previous Work}
\label{sec:harmony:compare}

Most of the the Haskell-based
work~\citep{magalhaes-harmtrace,koops-fharm,magalhaes-fcomp} has
focused on harmony. In particular \fharm~\citep{koops-fharm} tackles
the problem of harmonization of a melody, and their harmonization of
Piston's example is shown on the right side of
Figure~\ref{fig:harmonization}. Their method is quite different from
ours. Instead of using Piston's table of root progressions they use a
more sophisticated hierarchical grammar from previous
work~\citep{magalhaes-harmtrace}.

Their technique is to first require the melody notes to be members of
the harmonizing chords, then use their previous work to try to
create a harmonic analysis of every possible sequence of chords that
arises, choosing the one that most closely fits their
grammar. Although the result is not bad, it clearly leaves room for
improvement. Noted in the paper is that no attention is given to
voicing (how each of the four parts flows horizontally and how the
lines interact---in other words the counterpoint). Not noted is that
the melody note is always doubled, whereas typically in four-part
harmony it is preferred to double the root note. Also minimizing the
number of parse errors as a means to choose the best harmonization
does not necessarily have a musical meaning.

TODO: More comparison.