### 2020-03-04 doc/icfp20/conclusion.tex
final tweaks & swap figures

\section{Conclusion and Future Work}
\label{sec:conclusion}

% We believe that functional programming, and in particular functional
% programming with dependent types, is a promising way to do software
% engineering going forward. The languages and tools are perhaps not yet
% as finely polished and rich as those typically used in production
% software, but they are fundamentally far more powerful, and thus once
% the infrastructure (and training of programmers) catches up should
% become the languages of choice.

In this paper we have reported our experience in using Agda for
formalizing music theory.
We have found music to be a microcosm of issues that arise
everywhere in software engineering: modularity, composability,
correctness, and data issues such as handling equivalent or slightly
different formulations of the same data. If we can
demonstrate that functional programming with dependent types works
well in the domain of music, the same techniques can be used in
software in general.

Data issues are particularly prevalent at API boundaries between
systems, which may internally represent the data in different formats
(satisfying modularity and abstraction) but then rely on tedious and
potentially error-prone conversions between the formats. In music
we see this in wanting to work sometimes with absolute pitch and others
with relative pitch (a pair of an octave and a pitch within that
octave), or choosing between
a pair of pitches and a \texttt{PitchInterval}. Also we would like to
reuse a function that just works on a pitch to work on a note, which
consists of a pitch plus a duration.

For now we have explicitly converted between equivalent representations
and lifted functions, to get a feel for the amount of work
required.  However in the world of dependent types there are known
research techniques for handling these cases, namely transporting
across equivalences as in Homotopy Type Theory~\citep{hottbook} and
ornaments~\citep{dagand-ornaments}. Cubical
Agda~\citep{vezzosi-cubical}, although still early in development,
should be an excellent tool for exploring the extent to which these
research ideas can be applied in a practical context.

Aside from these more fundamental ideas there is simply the work of
further extending the formalization of music theory to encompass
harmonic analysis, more advanced counterpoint, voice leading, and even
composition~\citep{schoenberg-fundamentals}. The results so far have
been encouraging, but much more needs to be done to find the right
abstractions to express the rules in the simplest, clearest and most
composable way possible.
We expect the world of music theory to be worthy of a long exploration.