### 2020-04-21 doc/icfp20/response.txt
start response

We would like to thank the reviewers for their thorough reading and
valuable feedback.  The reviewers are right to point out our limited
use of dependent types; ...

Below is our answer to other questions.

### To Reviewer A

### To Reviewer B

* The best way to understand Line 315 is to look at the example in
  Figure 4.  In bar 5, the counterpoint line moves downwards from
  C to E (unison).  Then, in the next bar, it moves upwards from E to
  F.  We call a motion "step-wise" when its endpoints are adjacent
  notes, such as E-F.

### To Reviewer C

* The word "unspecified" in Line 87 is not a typo.  The unit of time is
  specified when the music is played at a particular tempo.