### 2020-04-23 doc/icfp20/response.txt
fix typo; address one more minor question

We would like to thank the reviewers for their thorough reading and
valuable feedback. We agree with the reviewers that the use of
dependent types in the paper is not sufficient to demonstrate their
true value in the context of music theory. There is additional use of
dependent types in the code that did not make it into the paper due to
space limitations and the difficulty of fitting it into the main
narrative, but likely this would not have been enough to change
anyone's mind. We have many more plans that should demonstrate how
compelling dependent types are in this context, but this is still
future work.

All reviewers raise other interesting points that we intend to
address, and we thank them for pointing these out.  Below we would
like to respond to the main points.

### Extrinsic vs. intrinsic approaches (Reviewers B and C)

The reviewers raise doubts about whether the extrinsic approach
in this paper is really superior to the intrinsic approach of previous
work. Although our experience so far has indicated the extrinsic
approach seems both more flexible and more user-friendly, a more
detailed comparison is warranted. We were unaware that Agda's
reflection mechanism allows custom error messages, so thanks to
Reviewer B for mentioning this and we will try it out. It is enticing
to take advantage of the built-in type checker as much as possible,
but it can also be constraining if one wants to for example explore
multiple rule systems at once. The two approaches are not so different
as indices can be represented as predicates attached to a $\Sigma$
(record) type, and with an appropriate equivalence can be converted
back and forth, which we also intend to explore. As Reviewer C notes
we need to be more clear about what advantages there are to running
the check at compile time rather than run-time in Haskell.

### Constraint-based search (Reviewers B and C)

The reviewers also mention constraint-based search in which the type
system may be used to partially constrain the search.  This is also
planned future work.

### Triads as a relation (Reviewer A)

The reviewer suggests implementing previousTriads as a relation instead
of a function. The tradeoff between these approaches, and indeed
many other aspects of practical programming using dependent types, are
areas we hope to explore in detail going forward.

[Note to Youyou: We could just leave out specific answers to questions
unless you think the reviewers are really interested in this.]

[Note to John: I agree that these questions are not very important,
but I would prefer not to ignore them.]

Lastly, let us answer some minor questions.

### To Reviewer B

* The reviewer asks how to interpret the sentence in Line 315.  The
  best way to understand it is to look at the example in Figure 4.
  In bar 5, the counterpoint line moves downwards from C to E (unison).
  Then, in the next bar, it moves upwards from E to F. We call a motion
  "step-wise" when its endpoints are adjacent notes, such as E-F.

* The reviewers wonders in what way harmony is dual to counterpoint.
  As Cong and Leo [FARM '19] suggest, we may view harmony as a
  horizontal view of a musical score, and counterpoint as a vertical view.

### To Reviewer C

* The word "unspecified" in Line 87 is not a typo. The unit of time is
  specified when the music is played at a particular tempo.